Iterative Reduction: Parsing and Interpreting Code Without Trees

### 2024-12-17 Paper/paper.tex
Responded to comments, added example

%\usepackage[margin=1in]{geometry}
\usepackage{amsmath, amssymb, mathtools}
\usepackage{enumitem}
\usepackage{float}
\usepackage{hyperref}
\usepackage{listings}
\usepackage{graphicx}
\usepackage{mdframed}
\usepackage{cleveref}

@@

\subsection{A First Attempt -- Primitive Reduction}
\label{sec:firstattempt}
We can formalize the algorithm given in the introduction as follows: we start with an alphabet of tokens $\Sigma$.
Define $\oS=\Sigma\times\bbN$, and we assume the existence of a so-called \textit{initial $\beta$-reducer}:
$\hb\colon \Sigma\times\Sigma \longto \Sigma$, which is defined by the implementer of the language being interpreted, and is the function which
tells us how to combine tokens together.
We extend it to a \textit{derived $\beta$-reducer} which is a function $\beta\colon\oS^*\longto\oS^*$ as follows.

@@

@@

\begin{example}
Define $\Sigma$ to consist of tokens of the form $\mbt{x}$ 
for every $x\in\mathbb{N}$, 
$\mb\circ$ for $\circ\in\{+,*\}$, and $\mbt{x,$\circ$}$
for every $x\in\mathbb{N}$
and $\circ\in\{+,*\}$.

@@

Now, consider the expression $1*2+3$. 
After assigning priorities, we obtain
the following expression:
\yoni{The last sentence is an example of how the assignment
of priorities feels a bit out of context here. This could improve
if they become a formal part of the framework, from the very beginning.}
$\mbt1_\infty \mbt*_2 \mbt2_\infty \mbt+_1 \mbt3_\infty$.
We now get:
\begin{multline*}
    \mbt1_\infty \mbt*_2 \mbt2_\infty \mbt+_1 \mbt3_\infty \to \mbt{1,*}_2 \mbt2_\infty \mbt+_1 \mbt3_\infty \to \mbt{1,*}_2 \mbt{2,+}_1 \mbt3_\infty \to \mbt{2,+}_1 \mbt3_\infty\\
    \to \mbt{2,+}_1 \mbt3_0 \to \mbt5_0
\end{multline*}
\end{example}

But this has a major issue: it can't handle things like parentheses.
We could of course handle parentheses by changing how we assign priorities: every time we see an open parentheses we could increase
the priority of each token and then beta-reduce.
But this complicates how we allocate priorities, 
\yoni{would u say that currently we can set priorities statically, but with
the idea of increasing priorities we would have to set them dynamically?
If so, I would mention that.}
and does not generalize well.
Furthermore the notion introduced in the next section, partial reduction, generalizes well and is used as a basepoint in the
remainder of this paper.

@@

\begin{example}
We define our alphabet $\Sigma$ to be the same as in~\Cref{sec:firstattempt}.
Now we define our initial $\beta$-reducer which will support parentheses.
It is an extension of our previous $\beta$-reducer:
\begin{gather*}
    \hb(\mbt x,\mb\circ) = (\mb{\ttt x,\circ},\snd),\quad \hb(\mb{\ttt x,\circ},\mbt y) = (\mb{\ttt x\circ\ttt y},\snd) \qquad \hbox{for $\circ\in\set{\ttt{+,*}}$} \\
    \hb(\mbt{x,*},\mbt{y,+}) = (\mbt{x*y,+},\snd)
\end{gather*}
where $\fst,\snd$ are the first and second projections, respectively (i.e. $(x,y)\mapsto x$ and $(x,y)\mapsto y$).
We then add the rules for parentheses:
\begin{gather*}
    \hb(\mbt x,\mbt)) = (\mbt{x,)},\snd),\quad \hb{\mb{\ttt x,\circ},\mbt{y,)}} = (\mb{\ttt x\circ\ttt y,\ttt)},\snd) \qquad \hbox{for $\circ\in\set{\ttt{+,*}}$} \\
    \hb(\mbt(,\mbt{x,)}) = (\mbt x,\fst)
\end{gather*}
\gobble) % syntax highlighting
And we define our priority function as follows:
\yoni{feels out of context, the notion of a priority function
was not defined. Also, notice that some tokens get priorities and some don't, which is a bit confusing.}
\[ \pi(\mbt x) = \infty\hbox{ for numbers \ttt x},\quad \pi(\mbt+) = 1,\quad \pi(\mbt*) = 2,\quad \pi(\mbt() = \infty,\quad \pi(\mbt)) = 0 \]

@@

Here we abstract away the value of each term; the value of the new term given by the initial $\beta$-reducer will be determined by an output function.
Our alphabet of types $\Sigma$, instead of having a token for every number, will just have tokens for types (number, operator, etc.).
We also have another set $\values$, the \textit{universe of values}.

\yoni{I would consider using the word "kinds" instead of "types". "kinds" is not great either, but types are stuff like int, real, int->int, which i am not sure is appropriate here.}

@@

\begin{example}
Define $\Sigma=\set{\NUM,\OP,\LPAREN,\RPAREN,\OP\NUM,\RPAREN\NUM}$.
Our initial $\beta$-reducer will is defined as follows:
\begin{itemize}
    \item $\hb(\NUM,\OP)=(\OP\NUM,\snd,(v,f\mapsto(v,f))$
    \item $\hb(\OP\NUM,\NUM)=(\NUM,\snd,((v,f),u\mapsto f(v,u)))$
    \item $\hb(\OP\NUM,\OP\NUM)=(\OP\NUM,\snd,((v,f),(u,g)\mapsto (f(v,u),g))$
    \item $\hb(\NUM,\RPAREN)=(\RPAREN\NUM,\snd,(v,\_\mapsto v))$
    \item $\hb(\OP\NUM,\RPAREN\NUM)=(\RPAREN\NUM,\snd,((v,f),u\mapsto f(v,u)))$
    \item $\hb(\LPAREN,\RPAREN\NUM)=(\NUM,\fst,(\_,v\mapsto v))$
\end{itemize}

\yoni{I don't really understand it. For example, what is 
$(v,f\mapsto(v,f))$? Perhaps introducing some more definitions
and details can help clarify?
}
\yoni{What happened to the squares around the tokens in this example?}

So taking our previous example of \ttt{2*(1+3)+4}: we first must convert it into tokens acceptable for $\beta$-reduction (via lexing).
So it is converted to
\[ \NUM_\infty(2)\ \OP_2(*)\ \LPAREN_\infty\ \NUM_\infty(1)\ \OP_1(+)\ \NUM_\infty(3)\ \RPAREN_0\ \OP_1(+)\ \NUM_\infty(4) \]

\yoni{Here we are missing a formal assignment of priorities.}

@@

We define the following four base sets:
\begin{enumerate}
    \item $\values$ the \textit{universe of values}; these are all the internal values an object may have.
    \item $\termp$ the set of \textit{printable terms}; this is the set of all printable tokens a programmer may type.
    \item $\terms$ the set of \textit{type terms}; these are terms which correspond to types in the classical sense.
    \item $\terma$ the set of \textit{abstract terms}; these are intermediate terms (like $\OP\NUM$ from the previous section).
\end{enumerate}
The sets $\termp,\terms,\terma$ are disjoint, and we place no such restriction on $\values$.
\yoni{The comment about no restrictions for $\values$ is unclear.}

While we don't need to place any further restrictions on $\termp,\terms,\terma$, we found it useful to explicitly construct them.
To construct $\terms$, we begin with a set of \textit{base types} $\Sigma$.
Each base type has an associated arity in $\bbN$; let $\Sigma^n$ denote the set of all base types of arity $n$.
Then we define $\terms$ as follows:
\[ \terms \ccoloneqq \Sigma^0 \mid \Sigma^1\terms \mid \Sigma^2\terms\terms \mid \Sigma^3\terms\terms\terms \mid \cdots \mid \Sigma^\oo(\terms\cdots\terms) \]

\yoni{I don't think this is very standard. Perhaps you can define
 it inductively: Every element of $\Sigma^0$ is in $T_{\Sigma}$,
  and for eac $n$, if $t_i\in T_{\Sigma}$ for each $1\leq i \leq n$
  then $f t_1 \ldots t_n \in T_{\Sigma}$ for each $f\in\Sigma^{n}$,
  and then for each $g\in\Sigma^{\infty}$ and any $n$, similarly...
Do we actually need $\Sigma^{\infty}$ BTW? }

To construct $\terma$ we start with $\mathcal A$, a set of \textit{atomic abstract terms}, again each with an arity in $\bbN$.
$\mathcal A^n$ is the set of atomic abstract terms of a set arity $n$.
Then $\terma$ is defined by
\[ \terma \ccoloneqq \mathcal A^0\mid\mathcal A^1\terms^\epsilon\mid\mathcal A^2\terms^\epsilon\terms^\epsilon\mid\mathcal A^3\terms^\epsilon\terms^\epsilon\terms^\epsilon\mid\cdots\mid
\mathcal A^\infty\terms^\epsilon\cdots\terms^\epsilon \]
Here we allow inputs to atomic abstract terms to be empty (i.e. $\terms^\epsilon$ instead of just $\terms$).

\begin{example}
If $\Sigma^{0}=\{\NUM\}$,
$\Sigma^{1}=\{\LIST\}$, and $\Sigma^{\infty}=\{\PRODUCT\}$
and $\Sigma=\Sigma^0\cup\Sigma^1\cup\Sigma^\infty$,
 and $\mathcal A=\set{\OP}$, 
\yoni{plz specify the "arity" of $\OP$.}
 then
\[ \NUM,\LIST\NUM,\PRODUCT(\LIST\NUM\NUM)\in\terms,\qquad \OP\LIST\NUM\in\terma \]
\yoni{I think it is better to put a space between these, especially
in $\LIST\NUM\NUM$.}
\end{example}
For readability we can put parentheses and commas where necessary, but importantly these sets have the \textit{unique reconstruction property}:
\yoni{If I understand correctly you are assuming this property, right? If so, please say so.}
it is possible to compute their construction from the
string itself (i.e. a string cannot be constructed from two different patterns).

Now we define the following sets:
\begin{enumerate}
    \item $\term=\termp\cup\terms\cup\terms$, 
\yoni{Should the last $\terms$ be $\terma$?}
    the set of \textit{terms},
    \item $\termi=\terms\cup\terma$, the set of \textit{internal terms},
    \item $\expi=\termi\times\values$, the set of \textit{internal expressions},
    \item $\exp=\termi\cup\termp$, the set of \textit{expressions}.
\end{enumerate}
Finally we define $\oe=\exp\times\bbZ$.
Elements of $\oe$ will be written like $\sigma_n(v)$ as before; 
\yoni{But elements of $\oe$ seems like pairs to me, not triplets.}
for printable terms we act as if the value is empty.
\yoni{Please clarify what does acting as if... means.}

The \textit{initial priority function} is a function $\pi\colon\termp\longto\bbZ$.
This can be naturally extended to $\pi\colon\termp^*\longto\oe^*$.
\yoni{please show how.}

Finally, our \textit{initial $\beta$-reducer} is a function ($\state$ is a set of functions $\termp\partial\expi$ which we define later)
\yoni{Using something and defining it later should be avoided...}

\yoni{Generally, this section is hard to follow. I suggest adding plenty of examples, after every new definition / concept, please add an appropriate small example.}

@@

>>>>

%\usepackage[margin=1in]{geometry}
\usepackage{amsmath, amssymb, mathtools}
\usepackage{enumitem}
\usepackage{float}
\usepackage{hyperref}
\usepackage{listings}
\usepackage{graphicx}
\usepackage{mdframed}
%\usepackage{cleveref}

@@

\subsection{A First Attempt -- Primitive Reduction} \label{sec:firstattempt}

We can formalize the algorithm given in the introduction as follows: we start with an alphabet of tokens $\Sigma$.
Define $\oS=\Sigma\times\bbN$, and we assume the existence of a so-called \textit{initial $\beta$-reducer}:
$\hb\colon \Sigma\times\Sigma \longto \Sigma$, which is defined by the implementer of the language being interpreted, and is the function which
tells us how to combine tokens together.
We extend it to a \textit{derived $\beta$-reducer} which is a function $\beta\colon\oS^*\longto\oS^*$ as follows.

@@

To assign priorities to an input stream of tokens, we assume the existence of a \textit{priority function} $\pi\colon\Sigma\partial\bbN$.
We can then extend this to a function $\pi\colon\Sigma^*\longto\oS^*$ by $\pi(\sigma^1\cdots\sigma^n)=\sigma^1_{\rho_1}\cdots\sigma^n_{\rho_n}$ where $\rho_i=\pi(\sigma^i)$.
Notice that not every token is assigned an initial priority; such tokens thus cannot be in the initial input stream to the $\beta$-reduction.

@@

\begin{example}

    Define $\Sigma$ to consist of tokens of the form $\mbt{x}$ for every $x\in\mathbb{N}$, $\mb\circ$ for $\circ\in\{+,*\}$, and $\mbt{x,$\circ$}$ for every $x\in\mathbb{N}$
    and $\circ\in\{+,*\}$.

@@

    We also define the initial priority function to give numbers a priority of $\infty$, and operations a priority corresponding to their precedence in the order of operations,
    where smaller priorities have less precedence.
    Namely, we give $+$ a priority of $1$ and $*$ a priority of $2$.

    Now, consider the expression $1*2+3$. 
    After assigning priorities, we obtain the following expression:
    $\mbt1_\infty \mbt*_2 \mbt2_\infty \mbt+_1 \mbt3_\infty$.
    \yoni{The last sentence is an example of how the assignment
    of priorities feels a bit out of context here. This could improve
    if they become a formal part of the framework, from the very beginning.}
    \ari{Done, thoughts?}
    We now get:
    \begin{multline*}
        \mbt1_\infty \mbt*_2 \mbt2_\infty \mbt+_1 \mbt3_\infty \to \mbt{1,*}_2 \mbt2_\infty \mbt+_1 \mbt3_\infty \to \mbt{1,*}_2 \mbt{2,+}_1 \mbt3_\infty \to \mbt{2,+}_1 \mbt3_\infty\\
        \to \mbt{2,+}_1 \mbt3_0 \to \mbt5_0
    \end{multline*}

But this has a major issue: it can't handle things like parentheses.
We could of course handle parentheses by changing how we assign priorities: every time we see an open parentheses we could increase
the priority of each token and then beta-reduce.
But this complicates how we allocate priorities, as we can no longer statically assign priorities using a simple priority function $\pi$ but must instead design a more complicated algorithm in order
to allow for generalization.
\yoni{would u say that currently we can set priorities statically, but with
the idea of increasing priorities we would have to set them dynamically?
If so, I would mention that.}
\ari{Done, thoughts?}
Furthermore the notion introduced in the next section, partial reduction, generalizes well and is used as a basepoint in the
remainder of this paper.

@@

\begin{example}

    We define our alphabet $\Sigma$ to be the same as in~\ref{sec:firstattempt}.
    Now we define our initial $\beta$-reducer which will support parentheses.
    It is an extension of our previous $\beta$-reducer:
    \begin{gather*}
        \hb(\mbt x,\mb\circ) = (\mb{\ttt x,\circ},\snd),\quad \hb(\mb{\ttt x,\circ},\mbt y) = (\mb{\ttt x\circ\ttt y},\snd) \qquad \hbox{for $\circ\in\set{\ttt{+,*}}$} \\
        \hb(\mbt{x,*},\mbt{y,+}) = (\mbt{x*y,+},\snd)
    \end{gather*}
    where $\fst,\snd$ are the first and second projections, respectively (i.e. $(x,y)\mapsto x$ and $(x,y)\mapsto y$).
    We then add the rules for parentheses:
    \begin{gather*}
        \hb(\mbt x,\mbt)) = (\mbt{x,)},\snd),\quad \hb{\mb{\ttt x,\circ},\mbt{y,)}} = (\mb{\ttt x\circ\ttt y,\ttt)},\snd) \qquad \hbox{for $\circ\in\set{\ttt{+,*}}$} \\
        \hb(\mbt(,\mbt{x,)}) = (\mbt x,\fst)
    \end{gather*}
    \gobble) % syntax highlighting
    And we define our priority function as follows:
    \yoni{feels out of context, the notion of a priority function
    was not defined. Also, notice that some tokens get priorities and some don't, which is a bit confusing.}
    \ari{Both points are now explained in the previous section, is this good?}
    \[ \pi(\mbt x) = \infty\hbox{ for numbers \ttt x},\quad \pi(\mbt+) = 1,\quad \pi(\mbt*) = 2,\quad \pi(\mbt() = \infty,\quad \pi(\mbt)) = 0 \]

@@

Here we abstract away the value of each term; the value of the new term given by the initial $\beta$-reducer will be determined by an output function.
Our alphabet of tokens $\Sigma$, instead of having a token for every number, will just have tokens for types (number, operator, etc.).
We also have another set $\values$, the \textit{universe of values}.

\yoni{I would consider using the word "kinds" instead of "types". "kinds" is not great either, but types are stuff like int, real, int->int, which i am not sure is appropriate here.}
\ari{I understand this, I'm not such a fan of ``kinds'' but I guess using ``types'' isn't really faithful to its understanding in CS.
I'd like to maybe discuss this with you more when we meet.}

@@

\begin{example}

    Define $\Sigma=\set{\NUM,\OP,\LPAREN,\RPAREN,\OP\NUM,\RPAREN\NUM}$.
    Our initial $\beta$-reducer will is defined as follows:
    \begin{itemize}
        \item $\hb(\NUM,\OP)=(\OP\NUM,\snd,(v,f\mapsto(v,f))$
        \item $\hb(\OP\NUM,\NUM)=(\NUM,\snd,((v,f),u\mapsto f(v,u)))$
        \item $\hb(\OP\NUM,\OP\NUM)=(\OP\NUM,\snd,((v,f),(u,g)\mapsto (f(v,u),g))$
        \item $\hb(\NUM,\RPAREN)=(\RPAREN\NUM,\snd,(v,\_\mapsto v))$
        \item $\hb(\OP\NUM,\RPAREN\NUM)=(\RPAREN\NUM,\snd,((v,f),u\mapsto f(v,u)))$
        \item $\hb(\LPAREN,\RPAREN\NUM)=(\NUM,\fst,(\_,v\mapsto v))$
    \end{itemize}
    In the first point $(v,f\mapsto(v,f))$ means that the function to compute the value of the new token is one which takes two values $v,f$ and outputs a new value which is the pair $(v,f)$.

    \yoni{I don't really understand it. For example, what is 
    $(v,f\mapsto(v,f))$? Perhaps introducing some more definitions
    and details can help clarify?
    }
    \ari{Explained, is this good?}
    \yoni{What happened to the squares around the tokens in this example?}

    So taking our previous example of \ttt{2*(1+3)+4}: we first must convert it into tokens acceptable for $\beta$-reduction (via lexing).
    In this example allocating initial priorities is not so simple: the priorities depend not on the type of the token in $\Sigma$, but the token in the stream given to the lexer.
    So we will assume the lexer allocates priorities similar to the previous example in the previous section.
    This issue will be resolved in the next section.
    So the stream is converted to
    \[ \NUM_\infty(2)\ \OP_2(*)\ \LPAREN_\infty\ \NUM_\infty(1)\ \OP_1(+)\ \NUM_\infty(3)\ \RPAREN_0\ \OP_1(+)\ \NUM_\infty(4) \]
    We no longer need boxes to distinguish between tokens because the individual tokens and so we will no longer use them.

    \yoni{Here we are missing a formal assignment of priorities.}
    \ari{I added a few sentences explaining this.}

@@

We define the following four base sets:
\begin{enumerate}
    \item $\values$ the \textit{universe of values}; these are all the internal values an object may have.
    \item $\termp$ the set of \textit{printable terms}; this is the set of all printable tokens a programmer may type.
    \item $\terms$ the set of \textit{type terms}; these are terms which correspond to types in the classical sense.
    \item $\terma$ the set of \textit{abstract terms}; these are intermediate terms (like $\OP\NUM$ from the previous section).
\end{enumerate}
The sets $\termp,\terms,\terma$ must be disjoint, and we place no such restriction on $\values$ (it can include elements from $\termp,\terms,\terma$).
\yoni{The comment about no restrictions for $\values$ is unclear.}

While we don't need to place any further restrictions on $\termp,\terms,\terma$, we found it useful to explicitly construct them.
To construct $\terms$, we begin with a set of \textit{base types} $\Sigma$.
Each base type has an associated arity in $\bbN$; let $\Sigma^n$ denote the set of all base types of arity $n$.
Then we define $\terms$ inductively as follows:
\begin{enumerate}
    \item for $\sigma\in\Sigma^0$; $\sigma$ is a type term.
    \item for $\sigma\in\Sigma^n$ for $n>0$, and $\tau_1,\dots,\tau_n$ type term; $\sigma\tau_1\cdots\tau_n$ is a type term.
    \item finally for $\sigma\in\Sigma^\infty$ and any $n\geq0$, if $\tau_1,\dots,\tau_n$ are type terms so is $\sigma(\tau_1\cdots\tau_n)$.
\end{enumerate}
We can write this definition as:
\[ \terms \ccoloneqq \Sigma^0 \mid \Sigma^1\terms \mid \Sigma^2\terms\terms \mid \Sigma^3\terms\terms\terms \mid \cdots \mid \Sigma^\oo(\terms\cdots\terms) \]

\yoni{I don't think this is very standard. Perhaps you can define
 it inductively: Every element of $\Sigma^0$ is in $T_{\Sigma}$,
  and for eac $n$, if $t_i\in T_{\Sigma}$ for each $1\leq i \leq n$
  then $f t_1 \ldots t_n \in T_{\Sigma}$ for each $f\in\Sigma^{n}$,
  and then for each $g\in\Sigma^{\infty}$ and any $n$, similarly...
Do we actually need $\Sigma^{\infty}$ BTW? }
\ari{Done.
And yeah, $\Sigma^\infty$ is important since products have a variable number of parameters.}

To construct $\terma$ we start with $\mathcal A$, a set of \textit{atomic abstract terms}, again each with an arity in $\bbN$.
$\mathcal A^n$ is the set of atomic abstract terms of a set arity $n$.
Then $\terma$ is defined as follows:
\begin{enumerate}
    \item for $\sigma\in\c A^0$; $\sigma$ is an abstract term.
    \item for $\sigma\in\c A^n$ for $n>0$, and $\tau_1,\dots,\tau_n\in\terms^\epsilon$; $\sigma(\tau_1,\dots,\tau_n)$ is an abstract term.
        When writing such an abstract term, we may omit parentheses and commas if there is no confusion.
    \item for $\sigma\in\c A^\infty$ and any $n\geq0$, for $\tau_1,\dots,\tau_n\in\terms^\epsilon$; $\sigma(\tau_1,\dots,\tau_n)$ is an abstract term.
\end{enumerate}
Alternatively:
\[ \terma \ccoloneqq \mathcal A^0\mid\mathcal A^1\terms^\epsilon\mid\mathcal A^2\terms^\epsilon\terms^\epsilon\mid\mathcal A^3\terms^\epsilon\terms^\epsilon\terms^\epsilon\mid\cdots\mid
\mathcal A^\infty\terms^\epsilon\cdots\terms^\epsilon \]
Here we allow parameters to atomic abstract terms to be empty (i.e. $\terms^\epsilon$ instead of just $\terms$).

For readability we can put and omit parentheses and commas where convenient, but importantly these sets have the \textit{unique reconstruction property}:
it is possible to compute their construction from the string itself (i.e. a string cannot be constructed from two different patterns).
This property is not hard to prove.
\yoni{If I understand correctly you are assuming this property, right? If so, please say so.}
\ari{No, by the definition of these sets, they have this property.}

\begin{example}

    If $\Sigma^{0}=\{\NUM\}$, $\Sigma^{1}=\{\LIST\}$, and $\Sigma^{\infty}=\{\PRODUCT\}$, and $\Sigma=\Sigma^0\cup\Sigma^1\cup\Sigma^\infty$, and $\c A=\mathcal A^1=\set{\OP}$, 
    \yoni{plz specify the "arity" of $\OP$.}
    \ari{Done}
    then
    \[ \NUM,\LIST\ \NUM,\PRODUCT(\LIST\ \NUM\ \NUM)\in\terms,\qquad \OP\ \LIST\ \NUM\in\terma \]
    \yoni{I think it is better to put a space between these, especially
    in $\LIST\NUM\NUM$.}
    \ari{Done}

Now we define the following sets:
\begin{enumerate}
    \item $\term=\termp\cup\terms\cup\terma$, 
        \yoni{Should the last $\terms$ be $\terma$?}
        \ari{Yeah}
    the set of \textit{terms},
    \item $\termi=\terms\cup\terma$, the set of \textit{internal terms},
    \item $\expi=\termi\times\values$, the set of \textit{internal expressions},
    \item $\exp=\expi\cup\termp$, the set of \textit{expressions}.
\end{enumerate}
Finally we define $\oe=\exp\times\bbZ$.
Elements of $\oe$ will be written like $\sigma_n(v)$ as before: $\sigma$ is the term (printable or internal), $n$ the priority, and $v$ the value.
In the case that $\sigma$ is a printable term, we still use this notation even though the value does not exist: so we just take it to be empty.
\yoni{But elements of $\oe$ seems like pairs to me, not triplets.}
\ari{There was a slight mistake in the definition of $\exp$.}
% for printable terms we act as if the value is empty.
\yoni{Please clarify what does acting as if... means.}
\ari{Rewritten}

The \textit{initial priority function} is a function $\pi\colon\termp\longto\bbZ$.
This can be naturally extended to $\pi\colon\termp^*\longto\oe^*$ as before: $\pi(\sigma^1\cdots\sigma^n)=\sigma^1_{\rho_1}\cdots\sigma^n_{\rho_n}$ where $\rho_i=\pi(\sigma^i)$.
\yoni{please show how.}
\ari{Done}

Finally, our \textit{initial $\beta$-reducer} is a function ($\state$ is a set of functions $\termp\partial\expi$ which we define later)
\yoni{Using something and defining it later should be avoided...}
\ari{We can move the definition up, but I felt like it messed with the flow.}

\yoni{Generally, this section is hard to follow. I suggest adding plenty of examples, after every new definition / concept, please add an appropriate small example.}
\ari{I'm not really sure what kinds of examples to add; I added a big example at the end that we can potentially split up?
But I feel like we need all of the definitions to give a good example.}

@@

\ari{Adding the following example:}

\begin{example}

    Let $\Sigma^0=\set\NUM$, $\Sigma^1=\set\LIST$, $\Sigma=\Sigma^0\cup\Sigma^1$; and finally we have $\c A^0=\set{\COLON,\LBRACK,\RBRACK}$, $\c A^1=\set\OP$, $\c A=\c A^0\cup\c A^1$.
    Let our set of printable terms consists of numbers as well as $\set{\mbt+,\mbt*,\mbt;,\mbt[,\mbt]}$.
    For readability, we again return to putting boxes around printable terms (but only \textit{printable} terms).
    Our initial priority function is
    \[ \pi(\mbt n) = \infty,\ \pi(\mbt+) = 1,\ \pi(\mbt*) = 2,\ \pi(\mbt;) = 0,\ \pi(\mbt[) = 0,\ \pi(\mbt]) = 0 \]
    We can transform the following stream \ttt{[1 + 2; 3 * 4;]} (which should be interpreted as the list $(3,12)$) into the stream of expressions:
    \[ \mbt[_0\ \mbt1_\infty\ \mbt+_1\ \mbt2_\infty\ \mbt;_0\ \mbt3_\infty\ \mbt*_2\ \mbt4_\infty\ \mbt;_0\ \mbt]_0 \]

    Our initial state will map numbers $\mbt n$ to $\NUM(n)$; operations $\mbt{$\circ$}$ to $\OP(\circ)$; $\mbt;$ to $\COLON$; $\mbt[$ to $\LBRACK$; and $\mbt]$ to $\RBRACK$.

    Our initial $\beta$-reducer is defined as follows: (we will neglect the input state and output state and string in the output of the $\beta$-reducer as they are held constant and empty, respectively.
    That is, instead of $\hb(\dots)=(\dots,(v,u,\bar s\mapsto\dots,\epsilon,\bar s))$ we will just write $\hb(\dots)=(\dots,(v,u\mapsto\dots))$.)
    \begin{itemize}
        \item $\hb(\NUM,\OP)=(\OP\NUM,\snd,(v,f,\mapsto(v,f)))$
        \item $\hb(\OP\NUM,\NUM)=(\NUM,\snd,((v,f),u,\bar s\mapsto f(v,u)))$
        \item $\hb(\OP\NUM,\OP\NUM)=(\OP\NUM,\snd,((v,f),(u,g)\mapsto (f(v,u),g)))$
        \item $\hb(\sigma,\COLON)=(\sigma,\zero,(v,\_\mapsto v))$
        \item $\hb(\LBRACK,\sigma)=(\LBRACK\sigma,\fst,(\_,v\mapsto(v)))$
        \item $\hb(\LBRACK\sigma,\sigma)=(\LBRACK\sigma,\fst,(\ell,v\mapsto\ell::(v)))$
        \item $\hb(\LBRACK\sigma,\RBRACK)=(\LIST\sigma,\pinfty,(\ell,\_\mapsto\ell))$
    \end{itemize}

    We now begin the $\beta$-reduction: 
    \ialign to\hsize{\rlap{$#$}\tabskip=0pt plus 1fil&$\to#$\hfil\tabskip=\z@\cr
    \mbt[_0\ \mbt1_\infty\ \mbt+_1\ \mbt2_\infty\ \mbt;_0\ \mbt3_\infty\ \mbt*_2\ \mbt4_\infty\ \mbt;_0\ \mbt]_0\cr
        &\LBRACK_0\ \mbt1_\infty\ \mbt+_1\ \mbt2_\infty\ \mbt;_0\ \mbt3_\infty\ \mbt*_2\ \mbt4_\infty\ \mbt;_0\ \mbt]_0\cr
        &\LBRACK_0\ \NUM_\infty(1)\ \mbt+_1\ \mbt2_\infty\ \mbt;_0\ \mbt3_\infty\ \mbt*_2\ \mbt4_\infty\ \mbt;_0\ \mbt]_0\cr
        &\LBRACK_0\ \NUM_\infty(1)\ \OP_1(+)\ \mbt2_\infty\ \mbt;_0\ \mbt3_\infty\ \mbt*_2\ \mbt4_\infty\ \mbt;_0\ \mbt]_0\cr
        &\LBRACK_0\ \OP\NUM_1(1,+)\ \mbt2_\infty\ \mbt;_0\ \mbt3_\infty\ \mbt*_2\ \mbt4_\infty\ \mbt;_0\ \mbt]_0\cr
        &\LBRACK_0\ \OP\NUM_1(1,+)\ \NUM_\infty(2)\ \mbt;_0\ \mbt3_\infty\ \mbt*_2\ \mbt4_\infty\ \mbt;_0\ \mbt]_0\cr
        &\LBRACK_0\ \OP\NUM_1(1,+)\ \NUM_\infty(2)\ \COLON_0\ \mbt3_\infty\ \mbt*_2\ \mbt4_\infty\ \mbt;_0\ \mbt]_0\cr
        &\LBRACK_0\ \OP\NUM_1(1,+)\ \NUM_0(2)\ \mbt3_\infty\ \mbt*_2\ \mbt4_\infty\ \mbt;_0\ \mbt]_0\cr
        &\LBRACK_0\ \NUM_0(3)\ \mbt3_\infty\ \mbt*_2\ \mbt4_\infty\ \mbt;_0\ \mbt]_0\cr
        &\LBRACK\NUM_0((3))\ \mbt3_\infty\ \mbt*_2\ \mbt4_\infty\ \mbt;_0\ \mbt]_0\cr
        &\LBRACK\NUM_0((3))\ \NUM_\infty(3)\ \mbt*_2\ \mbt4_\infty\ \mbt;_0\ \mbt]_0\cr
        &\LBRACK\NUM_0((3))\ \NUM_\infty(3)\ \OP_2(*)\ \mbt4_\infty\ \mbt;_0\ \mbt]_0\cr
        &\LBRACK\NUM_0((3))\ \OP\NUM_2(3,*)\ \mbt4_\infty\ \mbt;_0\ \mbt]_0\cr
        &\LBRACK\NUM_0((3))\ \OP\NUM_2(3,*)\ \NUM_\infty(4)\ \mbt;_0\ \mbt]_0\cr
        &\LBRACK\NUM_0((3))\ \OP\NUM_2(3,*)\ \NUM_\infty(4)\ \COLON_0\ \mbt]_0\cr
        &\LBRACK\NUM_0((3))\ \OP\NUM_2(3,*)\ \NUM_0(4)\ \mbt]_0\cr
        &\LBRACK\NUM_0((3))\ \NUM_0(12)\ \mbt]_0\cr
        &\LBRACK\NUM_0((3,12))\ \mbt]_0\cr
        &\LBRACK\NUM_0((3,12))\ \RBRACK_0\cr
        &\LIST\NUM_0((3,12))\cr
    }
